---
title: "A comment 'Orientation domains: A mobile grid clustering algorithm with spherical corrections'"
author: ""
date: "15.12.2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
```

```{r}
library(dendextend)
```


```{r}
library(dplyr)
```

```{r}
coeff = 180 / pi 
```

We are reading the file:
```{r}
orientations <- read.csv(file = "DataSet.csv", sep = ";", dec="," , header = TRUE)
```


```{r}
orientations$X_D<-cos(orientations$Dip_ang_deg / coeff) * cos( orientations$Dip_dir_deg / coeff);
orientations$Y_D<-cos(orientations$Dip_ang_deg / coeff) * sin(orientations$Dip_dir_deg / coeff)
orientations$Z_D<-sin(orientations$Dip_ang_deg / coeff)
```

Our data set has the following structure:
```{r}
head(orientations)
```


```{r}

```

We are sub-setting the data set by taking the direction cosines:
```{r}
subtable <- orientations[c("X_D", "Y_D", "Z_D")]
```

First, we perform the clustering using k-means algorithm:
```{r}
kmeans_r <- kmeans(subtable, centers = 3, iter.max = 50, nstart = 20)
```

The identification number of clusters are extracted:
```{r}
kmeansid <- kmeans_r$cluster
```

We summarise the clustering results using k-means algorithm:
```{r}
table(kmeansid)
```


We add the identification numbers to the initial data set:
```{r}
orientations <- mutate (orientations, kmeansid)
```

Secondly, we perform hierarchical clustering using euclidean distance and Ward linkage method:

```{r}
euclid_r <- hclust (dist(subtable, method="euclidean"), method="ward.D")
```

The identification number of clusters are extracted:

```{r}
euclidid <- cutree(euclid_r, k=3)
```

We summarise the clustering results using hierarchical clustering:
```{r}
table(euclidid)
```


```{r}
ggplot(orientations, aes(x=Dip_dir_deg, y=Dip_ang_deg))+
  geom_point()
```


We can add a dendrogram:

```{r}
dendroplot <- as.dendrogram(euclid_r)
```

We delete labels:
```{r}
labels(dendroplot) <- rep(c(""))
```

```{r}
tiff("Dendrogram.tif", units="in", width=7, height=5, res=1000)
dendroplot %>%
  set("branches_k_color", value = c(3,2,4), k = 3) %>% 
    plot(main = "Euclidean distance, \n Ward linkage", ylab="Distance")
dev.off()
```

We add the identification numbers to the initial data set:
```{r}
orientations <- mutate (orientations, euclidid)
```


The modified file looks as follows:

```{r}
head(orientations)
```
Saving data from different clusters to different files to upload in Stereonet.

```{r}
tab1<-dplyr::select(dplyr::filter(orientations, euclidid==1),c("Dip_ang_deg", "Dip_dir_deg"))
write.table(x= tab1, file = "DataSettab1.csv",  sep = ",", dec=".", col.names = TRUE, row.names = FALSE)
```

```{r}
tab2<-dplyr::select(dplyr::filter(orientations, euclidid==2),c("Dip_ang_deg", "Dip_dir_deg"))
write.table(x= tab2, file = "DataSettab2.csv",  sep = ",", dec=".", col.names = TRUE, row.names = FALSE)
```

```{r}
tab3<-dplyr::select(dplyr::filter(orientations, euclidid==3),c("Dip_ang_deg", "Dip_dir_deg"))
write.table(x= tab3, file = "DataSettab3.csv",  sep = ",", dec=".", col.names = TRUE, row.names = FALSE)
```

The modified file is saved:
```{r}
write.table(x= orientations, file = "DataSetResults.csv",  sep = ";", dec=".", col.names = TRUE, row.names = FALSE)
```

